### 2019-04-23 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
Talk to Emilie

Similarly to what we present above in Chapter~\ref{chapter:2}, Section~\ref{sec:2:lowerUpperBoundsRegret},
we use the Kullback-Leibler divergence to express our lower bound.
Let $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$ be KL divergence between the Bernoulli distribution of mean $x$ and that of mean $y$.
%
We first introduce the assumption under which we derive a regret lower bound, that generalizes a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.
Our lower bound will only be valid for algorithms of a certain class, that do not obtain much information from the collision indicator variables.
This class contains our proposals, \RandTopM{} and \MCTopM, but rule out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.

@@

\paragraph{The \RandTopM{} algorithm.}
%
First, we precisely state the \RandTopM{} algorithm above in Algorithm~\ref{algo:5:RandTopM}.
It is essentially a refinement over the \RhoRand{} algorithm, and a simpler version of \MCTopM.
The difference with \MCTopM{} is that the later algorithm introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
%
In \RandTopM{}, player $j$ is always considered ``not fixed'',
and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$.

@@

@@

\begin{figure}[h!]
  \resizebox{1.00\textwidth}{!}{
  \begin{tikzpicture}[>=latex',line join=bevel,scale=5]
      %
      \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
      \node (notfixed) at (1,0) [draw,rectangle,thick] {Not fixed, $\overline{s^j(t)}$};
      \node (fixed) at (0,0) [draw,rectangle,thick] {Fixed, $s^j(t)$};
      %
      \draw [black,->] (start) -> (notfixed.20);
      \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
      \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t), A^j(t) \in \TopM(t)$} (1);
      \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
      \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
      \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
      %
  \end{tikzpicture}
  }
  \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
  Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
  \label{fig:5:StateMachineAlgorithm_MCTopM}
\end{figure}

% -----------------------------------------------------------------
\section{Theoretical elements, and order optimal regret for \MCTopM}
\label{sec:5:upperbounds}
% -----------------------------------------------------------------

@@

The first experiment is considering $K=9$ arms, with means $\bm{\mu}=[0.1,0.2,\dots,0.9]$, and $M=3$ then $6$ then $9$ players.
Note that here and as in all this section we consider different algorithms that know the number of players $M$ (see Section~\ref{sub:5:unknownNumberOfPlayers} below for a discussion on the case when $M$ could be unknown).
%
Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance.
%
We report in Table~\ref{table:5:comparisonUCB_klUCB} the results in terms of mean regret, $\pm$ one standard deviation.
The conclusions are three fold: first we observe for any of the compared algorithms, using \klUCB{} is always (much) better than using \UCB.
Second we observe that our proposal \MCTopM{} is outperforming the state-of-the-art \RhoRand{} policy, and performs closely to the centralized (unrealistic) approach.
Third we observe in the last column, when $M=K=9$, that \MCTopM{} is achieving constant regret, as we proved (the regret upper-bound of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is $\cO(1)$ if there is no arms in \Mworst{} !), and it shows that the orthogonalization scheme proposed for \MCTopM{} is very efficient, especially in comparison to previous (state-of-the-art) approaches using ranks.
%
Additional experiments and illustrations are given below.
We illustrate below the (empirical mean) regret $R_t$ as a function of time and not only results in terms of empirical mean regret $R_T$ at the end of the bandit game.

>>>>

Similarly to what we present above in Chapter~\ref{chapter:2}, Section~\ref{sec:2:lowerUpperBoundsRegret},
we use the Kullback-Leibler divergence to express our lower bound.
Let $\kl(x,y) := x\log(x/y) + (1-x)\log((1-x)/(1-y))$ be KL divergence between the Bernoulli distribution of mean $x$ and that of mean $y$.
%
We first introduce the assumption under which we derive a regret lower bound, that generalizes a classical assumption made by \cite{LaiRobbins85} in single-player bandit models.
Our lower bound will only be valid for algorithms of a certain class, that do not obtain much information from the collision indicator variables.
This class contains our proposals, \RandTopM{} and \MCTopM, but rules out the \textsc{Sic-MMAB} algorithm of \cite{BoursierPerchet18}.

@@

@@

\paragraph{The \RandTopM{} algorithm.}
%
First, we precisely state the \RandTopM{} algorithm above in Algorithm~\ref{algo:5:RandTopM}.
It is essentially a refinement over the \RhoRand{} algorithm, to not use the indirection of ranks, and a simpler version of \MCTopM.
The difference with \MCTopM{} is that the later algorithm introduces a concept of a ``Chair'', by considering a binary ``being fixed'' state $s^j(t)$, as presented in Algorithm~\ref{algo:5:MCTopM} below.
%
In \RandTopM{}, player $j$ is always considered ``not fixed'',
and a \emph{collision always forces a uniform sampling of the next arm} from $\TopM(t)$.

@@

\begin{figure}[h!]
  \resizebox{1.00\textwidth}{!}{
  \begin{tikzpicture}[>=latex',line join=bevel,scale=5.5]
      %
      \node (start) at (1.5,0.30) {$(0)$ Start $t=0$};
      \node (notfixed) at (1,0) [draw,rectangle,thick] {Not fixed, $\overline{s^j(t)}$};
      \node (fixed) at (0,0) [draw,rectangle,thick] {Fixed, $s^j(t)$};
      %
      \draw [black,->] (start) -> (notfixed.20);
      \draw [color=cyan,thick,->] (notfixed) to[bend right] node[midway,above,text width=5cm,text centered,black] {\small $(1)$ $\overline{C^j(t)}, A^j(t) \in \TopM(t)$} (fixed);
      \path [color=blue,thick,->] (notfixed) edge[loop right] node[right,text width=4cm,text badly centered,black] {\small $(2)$  $C^j(t), A^j(t) \in \TopM(t)$} (1);
      \path [color=red,thick,->] (notfixed) edge[loop below] node[below,text centered,black] {\small $(3)$  $A^j(t) \notin \TopM(t)$} (1);
      \path [color=darkgreen,thick,->] (fixed) edge[loop left] node[left,text width=2.9cm,text badly centered,black] {\small $(4)$ $A^j(t) \in \TopM(t)$} (fixed);
      \draw [color=red,thick,->] (fixed) to[bend right] node[midway,below,text centered,black] {\small $(5)$  $A^j(t) \notin \TopM(t)$} (notfixed);
      %
  \end{tikzpicture}
  }
  \caption[``State-machine'' representation of \MCTopM]{Player $j$ using \MCTopM, represented as ``state machine'' with $5$ transitions.
  Taking one of the five transitions means playing one round of the Algorithm~\ref{algo:5:MCTopM}, to decide $A^j(t+1)$ using information of previous steps.}
  \label{fig:5:StateMachineAlgorithm_MCTopM}
\end{figure}

% -----------------------------------------------------------------
\section{Theoretical elements, and regret upper bound for \MCTopM}
\label{sec:5:upperbounds}
% -----------------------------------------------------------------

@@

The first experiment is considering $K=9$ arms, with means $\bm{\mu}=[0.1,0.2,\dots,0.9]$, and $M=3$ then $6$ then $9$ players.
Note that here and as in all this section we consider different algorithms that know the number of players $M$ (see Section~\ref{sub:5:unknownNumberOfPlayers} below for a discussion on the case when $M$ could be unknown).
%
Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance.

We report in Table~\ref{table:5:comparisonUCB_klUCB} the results in terms of mean regret, $\pm$ one standard deviation.
The conclusions are three fold: first we observe for any of the compared algorithms, using \klUCB{} is always (much) better than using \UCB.
Second we observe that our proposal \MCTopM{} is outperforming the state-of-the-art \RhoRand{} policy, and performs closely to the centralized (unrealistic) approach.
Third we observe in the last column, when $M=K=9$, that \MCTopM{} is achieving constant regret, as we proved (the regret upper-bound of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is $\cO(1)$ if there is no arms in \Mworst{} !), and it shows that the orthogonalization scheme proposed for \MCTopM{} is very efficient, especially in comparison to previous (state-of-the-art) approaches using ranks.
%
Additional experiments and illustrations are given below.
We illustrate below the (empirical mean) regret $R_t$ as a function of time and not only results in terms of empirical mean regret $R_T$ at the end of the bandit game.

### 2019-04-23 2-Chapters/5-Chapter/chapter5.tex
Talk to Emilie

We consider a bandit problem with $K=9$ arms, of means $0.1,0.2,\dots,0.8,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
As before, we include the centralized multiple-play \klUCB{} as an unachievably efficient baseline, the \Selfish-\klUCB{} algorithm as an heuristic which does not require the know $M$.
Then we compare the \RhoRand, \RandTopM{} and \MCTopM{} algorithms, using \klUCB, which know $M$ beforehand, with their extensions implementing the same algorithm as \rhoRandEst{} to estimate $M$ on the fly.

@@

To illustrate the difficulty of the ``no sensing'' case, we illustrate here the performances of different algorithms designed specifically for this setting.
As above, we consider a bandit problem with $K=9$ arms, of means $0.1,0.2,\dots,0.8,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
We include the \Selfish-\klUCB{} algorithm as an heuristic, as well as the Improved \MusicalChair{} algorithm from \cite{LugosiMehrabian18} and \textsc{Sic-MMAB} from \cite{BoursierPerchet18}.
The centralized multiple-play \klUCB{} is included as an unrealistic efficient baseline, but it is interesting to also add it here, as it does not use the sensing information but only the joint information (the reward) $r^j(t)$, as it directly affects the player in an orthogonal configuration and thus never encounters any collision (thus $r^j(t)=Y_{t,A^j(t)}$ for each $j,t$).
%
For the two other algorithms, we use the parameters advised in all papers: for \textsc{Sic-MMAB}, we used $T_0 = \lceil K \log(T) \rceil$ as advised by the authors.
For Improved-MC, we used $c=1$ which gives $g=235$, whereas in the paper the authors use $c=128$ for their analysis. We found empirically no difference when using different values of the constant $c$ or $g$, and unfortunately the regret of Improved-MC was always found to be linear\footnote{For instance Improved-MC obtained a mean regret $12150$ for $T=10000$, for $M=3,K=9$, but seeing one value for one horizon does not mean anything, however we also experimented with larger values of $T$ and found a same linear behavior.}.

@@

In all the models of IoT networks considered in this thesis, we assume that \emph{the different IoT devices cannot communicate with each other}, and can only communicate with a unique gateway.
Of course, this hypothesis can be removed, and in practice in some families of wireless networks, communication between players are possible.
Note that this extension is not considering a graph of distributed agents all playing cooperatively to solve a unique bandit game, like it is studied in the ``graph bandit'' problem.
What we are interested here is an extension of our model where players can send some bits to one or all the other players, at every time steps or some time steps.

@@

An interesting extension is thus to limit the communication between players, either to a small number of bits, or just one bit, at every or only some time steps.
We identified that the state-of-the-art on this direction of research consists in the two very recent works published in April 2019, while I was finishing this chapter: \cite{tao2019collaborative} and \cite{wang2019distributed}.
Distributed pure exploration is studied in \cite{tao2019collaborative}, where the proposed new lower bounds for the regret of any algorithm for the distributed best arm identification problem, under the fixed time or fixed confidence settings.
The also propose effective algorithms that asymptotically match their lower-bounds (up-to logarithmic factors, in some cases).
The second work studies distributed learning for (not necessarily stochastic) multi-armed bandits as well as linear bandits \cite{tao2019collaborative}.
For a distributed $K$-armed bandit with $M$ agents, they developed two protocols achieving near-optimal regret $\cO(\sqrt{M K T \log(T)})$,
and requiring little communication cost,
one is independent of the time horizon $T$ and use $\cO(M \log(T))$ communications,
and the other is independent of the number of arms $K$ and use $\cO(M K \log(M))$ communications.
Their model and algorithms fit in the distributed, decentralized framework we advertise in all this thesis, and this last work is very interesting.

>>>>

We consider a bandit problem with $K=9$ arms, of means $0.1,\dots,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
As before, we include the centralized multiple-play \klUCB{} as an unachievably efficient baseline, the \Selfish-\klUCB{} algorithm as an heuristic which does not require the know $M$.
Then we compare the \RhoRand, \RandTopM{} and \MCTopM{} algorithms, using \klUCB, which know $M$ beforehand, with their extensions implementing the same algorithm as \rhoRandEst, to estimate $M$ on the fly.

@@

To illustrate the difficulty of the ``no sensing'' case, we illustrate here the performances of different algorithms designed specifically for this setting.
As above, we consider a bandit problem with $K=9$ arms, of means $0.1,\dots,0.9$, and three different cases of $M=3,6,9$ players, for $100$ independent repetitions and an horizon of $T=10000$.
We include the \Selfish-\klUCB{} algorithm as an heuristic, as well as the Improved \MusicalChair{} algorithm from \cite{LugosiMehrabian18} and \textsc{Sic-MMAB} from \cite{BoursierPerchet18}.
The centralized multiple-play \klUCB{} is included as an unrealistic efficient baseline, but it is interesting to also add it here, as it does not use the sensing information but only the joint information (the reward) $r^j(t)$, as it directly affects the player in an orthogonal configuration and thus never encounters any collision (thus $r^j(t)=Y_{t,A^j(t)}$ for each $j,t$).
%
For the two other algorithms, we use the parameters advised in all papers: for \textsc{Sic-MMAB}, we used $T_0 = \lceil K \log(T) \rceil$ as advised by the authors.
For Improved-MC, we used $c=1$ which gives $g=235$, whereas in the paper the authors use $c=128$ for their analysis. We found empirically no difference when using different values of the constant $c$ or $g$, and unfortunately the regret of Improved-MC was always found to be linear\footnote{For instance Improved-MC obtained a mean regret $12150$ for $T=10000$, for $M=3,K=9$, but seeing one value for one horizon does not mean anything, however we also experimented with larger values of $T$ and found a same linear behavior.}.

@@

In all the models of IoT networks considered in this thesis, we assume that \emph{the different IoT devices cannot communicate with each other}, and can only communicate with a unique gateway.
Of course, this hypothesis can be removed, and in practice in some families of wireless networks, communication between devices are possible.
Note that this extension is not considering a graph of distributed agents all playing cooperatively to solve a unique bandit game, like it is studied in the ``graph bandit'' problem.
What we are interested here is an extension of our model where players can send some bits to one or all the other players, at every time steps or some time steps.

@@

An interesting extension is thus to limit the communication between players, either to a small number of bits, or just one bit, at every or only some time steps.
We identified that the state-of-the-art on this direction of research consists in the two very recent works
% published in April 2019, while I was finishing this chapter:
\cite{tao2019collaborative} and \cite{wang2019distributed}.
Distributed pure exploration is studied in \cite{tao2019collaborative}, where the proposed new lower bounds for the regret of any algorithm for the distributed best arm identification problem, under the fixed time or fixed confidence settings.
The also propose effective algorithms that asymptotically match their lower-bounds (up-to logarithmic factors, in some cases).
The second work studies distributed learning for (not necessarily stochastic) multi-armed bandits as well as linear bandits \cite{tao2019collaborative}.
For a distributed $K$-armed bandit with $M$ agents, they developed two protocols achieving near-optimal regret $\cO(\sqrt{M K T \log(T)})$,
and requiring little communication cost,
one is independent of the time horizon $T$ and use $\cO(M \log(T))$ communications,
and the other is independent of the number of arms $K$ and use $\cO(M K \log(M))$ communications.
Their model and algorithms fit in the distributed, decentralized framework we advertise in all this thesis, and this last work is very interesting.

### 2019-04-23 3-Appendices/2-Appendix/appendix2.tex
So long
Talk to Emilie

%!TEX root = ../PhD_thesis__Lilian_Besson
% ******************************* Thesis Appendix B ****************************
% \chapter{About another publication: ``What Doubling-Trick Can and Can't Do for Multi-Armed Bandits''}
\chapter{About Doubling-Trick for Multi-Armed Bandits}
\label{app:2:DoublingTricks}

@@

\TODOL{
I cannot include it raw without more work on the maths, we have to write only the best versions of our results.

But I am interested in including this article in my thesis, mainly because GLRklUCB in Chapter 6 is not anytime!

I need to try DoublingTrick + GLRklUCB, at least experimentally!
For the theoretical analysis, it will probably be a huge failure.
}

>>>>

%!TEX root = ../PhD_thesis__Lilian_Besson
% ******************************* Thesis Appendix B ****************************
% \chapter{About another publication: ``What Doubling-Trick Can and Can't Do for Multi-Armed Bandits''}
\chapter{About Doubling Tricks for Multi-Armed Bandits}
\label{app:2:DoublingTricks}

@@

% \TODOL{
% I cannot include it raw without more work on the maths, we have to write only the best versions of our results.

% But I am interested in including this article in my thesis, mainly because GLRklUCB in Chapter 6 is not anytime!

% I need to try DoublingTrick + GLRklUCB, at least experimentally!
% For the theoretical analysis, it will probably be a huge failure.
% }

### 2019-04-23 5-Nomenclature/Nomenclature.tex
Talk to Emilie

% --------------------------------
% matches Greek Symbols        > G
\nomenclature[G]{$\nu_k$, $\nu_k(t)$, $\nu_k^j$}{Distribution of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$}
\nomenclature[G]{$\mu_k$, $\mu_k(t)$, $\mu_k^j$, $\mu_i^{(k)}$}{Mean of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$ in Chapter~\ref{chapter:5}, arm $i$ in the $k$-th stationary interval in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\mu^*$, $\mu_k^*$}{Mean of the optimal arm (\ie, $\mu^*=\max_k \mu_k$), and mean of the $k$-th best arm}
\nomenclature[G]{$\bm{\mu},\bm{\nu},\bm{\lambda},$}{Vector of means $(\mu_k) = \mu_1,\dots,\mu_K$, or vector of distributions $(\nu_k)_k$ or $(\lambda_k)_k$ (characterizing a problem)}
\nomenclature[G]{$\Delta$}{Usually denotes the gap in terms of means of arms, usually between the best and second best arms, in Chapters~\ref{chapter:2}, \ref{chapter:5}. In Chapter~\ref{chapter:4} in Algorithm~\ref{algo:43:UCBwithDelay} it denotes a delay. In Chapter~\ref{chapter:6}, it can denote different gaps}
\nomenclature[G]{$\Delta n$, $\Delta s$}{Parameters of numerical optimization tricks in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\delta$}{Usually denotes a lower-bound on the gap $\Delta$, known before hand by an algorithm, in Chapters~\ref{chapter:5}, \ref{chapter:6}. Also denote a confidence level of an algorithm in Section~\ref{sec:6:ChangePointDetector}}
\nomenclature[G]{$\Upsilon_T$}{Number of break-points in a piece-wise stationary MAB problem in Chapter~\ref{chapter:6}. $\NCi$ denotes the number of change-points on arm $i$, and $C_T = \sum_{i=1}^{K} \NCi$ the number of change-points on the arms}
\nomenclature[G]{$\tau_k^j$, $\tau^{(k)}$}{Usually denotes the location of a change-point, \eg, the $j$-th change-point on arm $k$, in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\tau$}{Usually denotes the length of a sliding-window algorithm like Sliding-Window-\UCB{} in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\gamma$}{Usually denotes the discount factor of an algorithm like Discounted-\UCB{} in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\varepsilon$}{Usually denotes a small positive real value, \eg, $\lambda_k = \mu_k + \varepsilon$ in Chapter~\ref{chapter:5} or a drift correction parameter in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\alpha$}{Usually denotes the parameter of a \UCB{} algorithm in Chapters~\ref{chapter:2} and \ref{chapter:4}, or a parameter controlling the forced exploration mechanism in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\beta$}{Usually denotes a threshold of GLR tests in Chapter~\ref{chapter:6}}

@@

% --------------------------------
% matches mathematical symbols > O
\nomenclature[O]{$\kl$}{Binary relative entropy, KullBack-Leibler divergence between two Bernoulli distribution: $\kl(x,y) = x\log(x/y) + (1-x) \log((1-x)/(1-y))$}
\nomenclature[O]{$\kl'$}{Derivative of the binary relative entropy function $\kl(x,y)$, with respect to $x$, for a fixed $y$, in Chapter~\ref{chapter:5}, see Lemma~\ref{lem:5:Fact2KLUCB}}
\nomenclature[O]{$d(x,y)$}{Divergence function between two distributions of parameters $x$ and $y$, in a one-dimensional exponential family, in Chapter~\ref{chapter:6}}

@@

% --------------------------------
% matches Subscripts           > I
\nomenclature[I]{$x_k$}{Usually the subscript index $k$ is denotes a variable depending on an arm, for $k\in\{1,\dots,K\}$}

% --------------------------------
% matches Superscripts         > E
\nomenclature[E]{$y^j$}{Usually the superscript index $j$ is denotes a variable depending on a player, for $j\in\{1,\dots,M\}$}
\nomenclature[E]{$y^{(k)}$, $y^{(\ell)}$}{Usually the superscript index $(k)$ (resp. $(\ell)$) is denotes a variable in the $k$-th stationary interval, under the \emph{global} (resp. \emph{local}) changes point-of-view, in Chapter~\ref{chapter:6}}

>>>>

% --------------------------------
% matches Greek Symbols        > G
\nomenclature[G]{$\nu_k$, $\nu_k(t)$, $\nu_k^j$}{Distribution of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$}
\nomenclature[G]{$\mu_k$, $\mu_k(t)$, $\mu_k^j$, $\mu_i^{(k)}$}{Mean of arm $k$, arm $k$ at time $t$, arm $k$ for user $j$ in Chapter~\ref{chapter:5}, arm $i$ in the $k$-th stationary interval in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\mu^*$, $\mu_k^*$}{Mean of the optimal arm (\ie, $\mu^*=\max_k \mu_k$), and mean of the $k$-th best arm}
\nomenclature[G]{$\bm{\mu},\bm{\nu},\bm{\lambda},$}{Vector of means $(\mu_k) = \mu_1,\dots,\mu_K$, or vector of distributions $(\nu_k)_k$ or $(\lambda_k)_k$ (characterizing a problem)}
\nomenclature[G]{$\Delta$}{Usually denotes the gap in terms of means of arms, usually between the best and second best arms, in Chapters~\ref{chapter:2}, \ref{chapter:5}. In Chapter~\ref{chapter:4} in Algorithm~\ref{algo:43:UCBwithDelay} it denotes a delay. In Chapter~\ref{chapter:6}, it can denote different gaps}
\nomenclature[G]{$\Delta n$, $\Delta s$}{Parameters of numerical optimization tricks in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\delta$}{Usually denotes a lower-bound on the gap $\Delta$, known before hand by an algorithm, in Chapters~\ref{chapter:5}, \ref{chapter:6}. Also denote a confidence level of an algorithm in Section~\ref{sec:6:ChangePointDetector}}
\nomenclature[G]{$\Upsilon_T$}{Number of break-points in a piece-wise stationary MAB problem in Chapter~\ref{chapter:6}. $\NCi$ denotes the number of change-points on arm $i$, and $C_T = \sum_{i=1}^{K} \NCi$ the number of change-points on the arms}
\nomenclature[G]{$\tau_k^j$, $\tau^{(k)}$}{Usually denotes the location of a change-point, \eg, the $j$-th change-point on arm $k$, in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\tau$}{Usually denotes the length of a sliding-window, \eg, in SW-\UCB{} in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\gamma$}{Usually denotes a discount factor, \eg, in D-\UCB{} or D-TS in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\varepsilon$}{Usually denotes a small positive real value, \eg, $\lambda_k = \mu_k + \varepsilon$ in Chapter~\ref{chapter:5} or a drift correction parameter in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\alpha$}{Usually denotes the parameter of a \UCB{} algorithm in Chapters~\ref{chapter:2} and \ref{chapter:4}, or a parameter controlling the forced exploration mechanism in Chapter~\ref{chapter:6}}
\nomenclature[G]{$\beta$}{Usually denotes a threshold of GLR tests in Chapter~\ref{chapter:6}}

@@

% --------------------------------
% matches mathematical symbols > O
\nomenclature[O]{$\kl$}{Binary relative entropy, KullBack-Leibler divergence between two Bernoulli distribution: $\kl(x,y) = x\log(x/y) + (1-x) \log((1-x)/(1-y))$}
% \nomenclature[O]{$\kl'$}{Derivative of the binary relative entropy function $\kl(x,y)$, with respect to $x$, for a fixed $y$, in Chapter~\ref{chapter:5}, see Lemma~\ref{lem:5:Fact2KLUCB}}
\nomenclature[O]{$d(x,y)$}{Divergence function between two distributions of parameters $x$ and $y$, in a one-dimensional exponential family, in Chapter~\ref{chapter:6}}

@@

% --------------------------------
% matches Subscripts           > I
\nomenclature[I]{$x_k$}{Usually denotes a variable depending on an arm, for $k\in\{1,\dots,K\}$}

% --------------------------------
% matches Superscripts         > E
\nomenclature[E]{$y^j$}{Usually denotes a variable depending on a player, for $j\in\{1,\dots,M\}$}
\nomenclature[E]{$y^{(k)}$, $y^{(\ell)}$}{Usually the superscript index $(k)$ (resp. $(\ell)$) denotes a variable in the $k$-th stationary interval, under the \emph{global} (resp. \emph{local}) changes point-of-view, in Chapter~\ref{chapter:6}}
